### 2016-03-15 paper/icfp16-submission/evaluation.tex
address various minor reviewer comments

\paragraph{Benchmarks}
We answer these questions on two sets of ill-typed programs, \ie
programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages course at our institution.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB \# hidden for blind review), % \#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml programs.
%
The second dataset --- widely used in the literature --- comes from a
similar course at the University of Washington~\cite{lerner_seminal:_2006},
from which we extracted 284 ill-typed programs.

@@

\paragraph{Results}
\label{sec:results-witness}
The results of our experiments are summarized in
Figure~\ref{fig:results-witness}.
%
In both datasets our tool was able to find a witness for 83\% of the
programs in under one second, \ie fast enough to be integrated as a
compile-time check. If we extend our tolerance to a 10 second timeout,
we hit a maximum of 87\% coverage.
%
Interestingly, while the vast majority of witnesses corresponded to a
type-error, as expected, 3-4\% triggered an unbound variable error and
2-3\% triggered an infinite recursion error.
%
For the remaining 12\% of programs we were unable to provide any useful
feedback as they either completed 1,000 tests successfully, or timed out
after 10 minutes.
%
% XX programs were deemed safe and XX timed out even at 3,000 steps, \ie
% we could not provide any useful feedback for XX\% of the total programs.
%
While a more advanced search procedure, \eg dynamic-symbolic execution,
could likely trigger more of the type errors, our experiments show that
type errors are coarse enough (or that novice programs are \emph{simple}
enough) that these techniques are not necessary.

>>>>

\paragraph{Benchmarks}
We answer these questions on two sets of ill-typed programs, \ie
programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages course at our institution.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB \# hidden for blind review), % \#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml programs.
%
The second dataset --- widely used in the literature --- comes from a
similar course at the University of Washington~\cite{lerner_seminal:_2006},
from which we extracted 284 ill-typed programs.
%
Both datasets include programs that demonstrate a variety of functional
programming idioms, including (tail) recursive functions, higher-order
functions, polymorphic data types, and expression evaluators.

@@

\paragraph{Results}
\label{sec:results-witness}
The results of our experiments are summarized in
Figure~\ref{fig:results-witness}.
%
In both datasets our tool was able to find a witness for 83\% of the
programs in under one second, \ie fast enough to be integrated as a
compile-time check. If we extend our tolerance to a 10 second timeout,
we hit a maximum of 87\% coverage.
%
Interestingly, while the vast majority of witnesses corresponded to a
type-error, as expected, 3-4\% triggered an unbound variable error (even
though \ocaml reported a type error) and 2-3\% triggered an infinite
recursion error.
%
For the remaining 12\% of programs we were unable to provide any useful
feedback as they either completed 1,000 tests successfully, or timed out
after 10 minutes.
%
% XX programs were deemed safe and XX timed out even at 3,000 steps, \ie
% we could not provide any useful feedback for XX\% of the total programs.
%
While a more advanced search procedure, \eg dynamic-symbolic execution,
could likely trigger more of the type errors, our experiments show that
type errors are coarse enough (or that novice programs are \emph{simple}
enough) that these techniques are not necessary.

### 2016-03-15 paper/icfp16-submission/introduction.tex
address various minor reviewer comments

\paragraph{2. Visualizing Witnesses}
Our second contribution is an interactive visualization of the
execution of \ocaml\ programs, well-typed or not~(\S~\ref{sec:interactive}).
%
We extend the semantics to also build a \emph{reduction graph}
which records all of the small-step reductions and the context
in which they occur.
%
The graph lets us build a visualization that shows a sequence of
steps from the source witness to the stuck term and allows the user
to interactively expand the computation to expose intermediate steps,
by selecting an expression and choosing a traversal strategy.
%
The strategies include many of the standard debugging moves, \eg\
stepping \emph{forward} or \emph{into} or \emph{over} calls, as well
stepping or jumping \emph{backward} to understand how a particular
value was created, while preserving a context of the intermediate
steps that allow the user to keep track of a term's provenance.

>>>>

\paragraph{2. Visualizing Witnesses}
Our second contribution is an interactive visualization of the
execution of purely functional \ocaml\ programs, well-typed or not~(\S~\ref{sec:interactive}).
%
We extend the semantics to also build a \emph{reduction graph}
which records all of the small-step reductions and the context
in which they occur.
%
The graph lets us build a visualization that shows a sequence of
steps from the source witness to the stuck term and allows the user
to interactively expand the computation to expose intermediate steps,
by selecting an expression and choosing a traversal strategy.
%
The strategies include many of the standard debugging moves, \eg\
stepping \emph{forward} or \emph{into} or \emph{over} calls, as well
stepping or jumping \emph{backward} to understand how a particular
value was created, while preserving a context of the intermediate
steps that allow the user to keep track of a term's provenance.

### 2016-03-15 paper/icfp16-submission/overview.tex
address various minor reviewer comments

Returning to @fac@, given a hole as input we will narrow the hole
to an @int@ upon reaching the @<=@ test.
%
At this point we choose a random @int@ for the instantiation and
concrete execution takes over entirely, leading us to the expected crash
in the multiplication.

@@

\paragraph{Problem: How many inputs does a function take?}
%
There is another wrinkle though, how did we know
that @fac@ takes a single argument instead of two
(or none)?
%
It is clear, syntactically, that @fac@ takes \emph{at least} one
argument, but in a higher-order language with currying, syntax can be
deceiving.
%
Consider the following definition:
%
\begin{code}
  let incAllByOne = List.map (+ 1)
\end{code}
%
Is @incAllByOne@ a function?
%
If so, how many arguments does it take?
%
The \ocaml\ compiler deduces that @incAllByOne@ takes a single argument
because the \emph{type} of \hbox{@List.map@} says it takes two arguments, and it is
partially applied to @(+ 1)@.
%
As we are dealing with ill-typed programs we do not have the luxury of
typing information.

>>>>

Returning to @fac@, given a hole as input we will narrow the hole
to an @int@ upon reaching the @<=@ test.
%
At this point we choose a
random @int@\footnote{With standard heuristics to favor small values.}
for the instantiation and
concrete execution takes over entirely, leading us to the expected crash
in the multiplication.

@@

\paragraph{Problem: How many inputs does a function take?}
%
There is another wrinkle, though; how did we know
that @fac@ takes a single argument instead of two
(or none)?
%
It is clear, syntactically, that @fac@ takes \emph{at least} one
argument, but in a higher-order language with currying, syntax can be
deceiving.
%
Consider the following definition:
%
\begin{code}
  let incAllByOne = List.map (+ 1)
\end{code}
%
Is @incAllByOne@ a function?
%
If so, how many arguments does it take?
%
The \ocaml\ compiler deduces that @incAllByOne@ takes a single argument
because the \emph{type} of \hbox{@List.map@} says it takes two arguments, and it is
partially applied to @(+ 1)@.
%
As we are dealing with ill-typed programs we do not have the luxury of
typing information.
